---
title: "Introduction to deconICA"
subtitile: "Deconvolution of transcriptome through Immune Component Analysis"
author: "Urszula Czerwinska"
date: "`r Sys.Date()`"
output: 
    prettydoc::html_pretty:
        theme: cayman
        highlight: vignette
vignette: >
  %\VignetteIndexEntry{"Introduction to deconICA"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here is an introduction to the [`deconICA`](https://github.com/UrszulaCzerwinska/DeconICA) R package. **DeconICA** stands for **Deconvolution of transcriptome through Immune Component Analysis**. The aim of the project is to adapt blind source separation techniques to extract immune-related signals from mixed biological samples. A great exaple of mixed biological sample is transcriptome measured in heterogenous tissue such as blood or tumor biopsy. 

In this vignette we present short introduction to the blind source spearation techniques, the biological foundation of the problem and finally we walk you through examples on how to use [`deconICA`](https://github.com/UrszulaCzerwinska/DeconICA) R package.

If you are intrested only in practical examples of [`deconICA`](https://github.com/UrszulaCzerwinska/DeconICA), skip directly to [Tutorial](#Tut) section.

# Background

## Blind source separation

### Independent Components Analysis

### NMF

### convex hull methods

### attractor metagenes

## Biological mixed samples

### transcriptome


# Tutorial {#Tut}

## How to install
You can install `deconICA` from GitHub with:
```{r, eval = FALSE, echo = TRUE}
#install.packages("devtools")
devtools::install_github("UrszulaCzerwinska/DeconICA", build_vignettes = TRUE)
```
or 
```{r, eval = FALSE, echo = TRUE}
install.packages("githubinstall")
githubinstall::githubinstall("DeconICA", build_vignettes = TRUE)
```

[**TO DO**] You can install the stable version from [CRAN](https://cran.r-project.org/web/packages/deconica/index.html)
```{r, eval = FALSE, echo = TRUE}
install.packages('deconica', dependencies = TRUE)
```

Then load package with
```{r}
library(deconica)
```

## Overview of functions

You can use `run_fastica()` function to decompose a matrix into independent components
```{r}
S <- matrix(runif(10000), 5000, 2)
A <- matrix(c(1, 1, -1, 3), 2, 2, byrow = TRUE)
X <- data.frame(S %*% A)
res <- run_fastica(X, row.center = TRUE, n.comp = 2, overdecompose = FALSE)
str(res)
```
In this trival example we create sources matrix `S`and mixing matrix `A` that we multuply to obtain `X`. Then we decompose `X` into `n.comp = 2`, with row centering (substracting mean from each row) `{r} row.center = TRUE`. We also checked `overdecompose = FALSE` as `overdecompose = TRUE`


## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
